---
title: "R Project 1"
output: 
  github_document: 
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Required Packages
```{r, echo=TRUE, message=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
```

# Contact the NHL records API and the NHL stats API  

## nhl_records function:
```{r}
nhl_records <- function(endpoint, team_id=NULL, team_name=NULL){
  # Endpoint options are: Franchise, TeamTotals, Franchise_SeasonTotals, GoalieRecords, SkaterRecords
  base_url <- 'https://records.nhl.com/site/api'
  if(endpoint == 'Franchise'){
    fork <- '/franchise' 
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url)
  }
  if(endpoint == 'TeamTotals'){
    fork <- '/franchise-team-totals'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url)
  }
  if(endpoint == 'Franchise_SeasonTotals'){
    fork <- '/franchise-season-records'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  if(endpoint == 'GoalieRecords'){
    fork <- '/franchise-goalie-records'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  if(endpoint == 'SkaterRecords'){
    fork <- '/franchise-skater-records' 
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  
  # Convert to text
  df_text <- content(df_get, as='text', enconding='UTF-8')
  
  # Convert form JSON
  df_json <- fromJSON(df_text, flatten=TRUE)
  
  # Use conditional logic to determine how the user is searching
  if(!is.null(team_name) & is.null(team_id)){
    if(endpoint=='Franchise'){
      df <- df_json$data[,-1]
      
      # Clean the dataframe to make the names searchable in our current format
      teamCommonName <- paste(df$teamPlaceName, df$teamCommonName)
      df <- cbind(df, teamCommonName)
    
      # Remove the old names split over two columns
      df <- df[,-(4:5)]
    
      df <- filter(df, teamCommonName==team_name)
    }
    if(endpoint=='TeamTotals'){
      df <- df_json$data[,-1]
      df <- filter(df, teamName==team_name)
    }
    if(endpoint=='Franchise_SeasonTotals' | endpoint=='GoalieRecords' | endpoint=='SkaterRecords'){
      df <- df_json$data[,-1]
      df <- filter(df, franchiseName==team_name)
    }
  }
  if(is.null(team_name) & !is.null(team_id)){
    if(endpoint=='Franchise'){
      df <- df_json$data[,-1]
      df <- filter(df, mostRecentTeamId==as.integer(team_id))
    }
    if(endpoint=='TeamTotals'){
      df <- df_json$data[,-1]
      df <- filter(df, teamId==team_id)
    }
    if(endpoint=='Franchise_SeasonTotals' | endpoint=='GoalieRecords' | endpoint=='SkaterRecords'){
      df <- df_json$data[,-1]
      df <- filter(df, franchiseId==team_id)
    }
  }  
  if(is.null(team_id) & is.null(team_name)){
    df <- df_json$data[,-1]
  }
  return(df)
}
```
```{r, echo=TRUE, eval=FALSE}
nhl_records(endpoint, team_id=NULL, team_name=NULL)
```
The `nhl_records` function pulls data from `records.nhl.com` and outputs it to a data frame. It takes on multiple arguments:  
  
* `endpoint`: can take on five values (`Franchise, TeamTotals, Franchise_SeasonTotals, GoalieRecords, SkaterRecords`). This option will pull data from different API tables. The value for this variable must be passed as a string.  
* `team_id`:  is a unique numeric identifier for each NHL team. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  
* `team_name`: is a unique character assignment for each NHL team. Team names take on the format '<City> <Club Name>'. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  

## nhl_stats function:
```{r}
nhl_stats <- function(modifier, modifier_extension=NULL, team_id=NULL){
  # Modifier options are: team_roster, person_names, team_schedule_next, team_schedule_previous, 
  # team_stats, team_roster_season, team_id, single_season_playoffs
  base_url <- paste0('https://statsapi.web.nhl.com/api/v1/teams/', team_id)
  
  if(modifier=='team_roster'){
    # Pull a GET request
    query <- paste0('?expand=team.roster', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='person_names'){
    # Pull a GET request
    query <- paste0('?expand=person.names', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_schedule_next'){
    # Pull a GET request
    query <- paste0('?expand=team.schedule.next', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_schedule_previous'){
    # Pull a GET request
    query <- paste0('?expand=team.schedule.previous', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_stats'){
    # Pull a GET request
    query <- paste0('?expand=team.stats', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_roster_season'){
    # Pull a GET request
    query <- paste0('?expand=team.roster&season=', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_id'){
    # Pull a GET request
    query <- paste0('?teamId=', modifier_extension)
    df_get <- GET(base_url, query=modifier_full)
  }
  if(modifier=='single_season_playoffs'){
    # Pull a GET request
    query <- paste0('?stats=statsSingleSeasonPlayoffs', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  
  # Convert to text
  df_text <- content(df_get, as='text', enconding='UTF-8')
  
  # Convert form JSON
  df_json <- fromJSON(df_text, flatten=TRUE)
  
  # Remove the redundant ID column
  df <- df_json$teams[,-1]
  
  return(df)
}
```
```{r, echo=TRUE, eval=FALSE}
nhl_stats(modifier, modifier_extension=NULL, team_id=NULL)
```
The `nhl_stats` function pulls data from `www.statsapi.web.nhl.com` and outputs it to a data frame. It takes on multiple arguments:  
  
* `modifier`: can take on one of eight values (`team_roster, person_names, team_schedule_next, team_schedule_previous, team_stats, team_roster_season, team_id, single_season_playoff`). This value will run an expand command on the API. The value for this variable must be passed as a string.  
* `modifier_extension`: Allows the user to pass additional search options to the API. An example of this would be `&gameType=P` to pull data from playoff games only. By default, this value is set to `NULL`. The value for this variable must be passed as a string.  
* `team_id`: is a unique numeric identifier for each NHL team. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  
  
## nhl_data function:
```{r}
nhl_data <- function(source, ...){
  if(source == 'records'){
    return(nhl_records(...))
  }
  if(source == 'stats'){
    return(nhl_stats(...))
  }
}
```
```{r, eval=FALSE, echo=TRUE}
nhl_data(source, ...)
```
The `nhl_data` function serves as a wrapper function for `nhl_records` and `nhl_stats`. 
  
  * `source`: calls whichever function is needed (`records` or `stats`). This argument must be passed a string.  
  * `...`: allows the user to pass all arguments supported by the functions `nhl_records` and `nhl_stats`.

# Some exploratory data analysis  

## Do a join on two returned datasets from different API endpoints
```{r, echo=TRUE, message=FALSE}
# Import the first data set using 'nhl_records' via the wrapper function
dat_1 <- nhl_data('records', endpoint='Franchise', team_name='Boston Bruins')

# Import the second data set using 'nhl_stats' via the wrapper function 
dat_2 <- nhl_data('stats', modifier='team_stats', team_id=6)

# Rename the full team name variable 'name' in 'dat_2' to prepare to merge the data sets
dat_2 <- dplyr::rename(dat_2, teamCommonName=name)

# Merge the two datasets
dat_3 <- merge(dat_1, dat_2, by='teamCommonName')

# Partial output
dat_3[1:5]
```

## Create at least two new variables  

* New variable: __Average games started__  
This variable will look at the average amount of games started by Carolina Hurricane goaltenders.
```{r, echo=TRUE, message=FALSE}
# Create a new variable, average_games_started
car_goal <- nhl_data('records', endpoint='GoalieRecords', team_name='Carolina Hurricanes')
car_avg_game <- mean(car_goal$gamesPlayed)
average_games_started <- rep(car_avg_game, times=nrow(car_goal))

# Add a new column corresponding to the average number of games a goaltender starts as a Hurricane
car_goal <- cbind(car_goal, average_games_started)

# Display a partial output
head(car_goal[,c(2,7,4,6,29)])
```

* New variable: __Hat trick__  
A hat trick is three or more goals scored by a player in a single game. This variable will determine if a player has scored a hat trick in their career.
```{r, echo=TRUE, message=FALSE}
# Create a new variable, hat_trick
bos_skate <- nhl_data('records', endpoint='SkaterRecords', team_name='Boston Bruins')
hat_trick <- ifelse(bos_skate$mostGoalsOneGame > 2, TRUE, FALSE)

# Add a column corresponding to the new variable hat_trick
bos_skate <- cbind(bos_skate, hat_trick)

# A partial output including our new variable
head(bos_skate[,c(3,9,5,8,15,30)])
```

## Create some contingency tables
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Import the entire 'SkaterRecords' data frame for the original six teams
original_six <- c('Chicago Blackhawks', 'Boston Bruins', 'Montréal Canadiens', 'Detroit Red Wings', 
                  'New York Rangers', 'Toronto Maple Leafs')
origsix_skate <- nhl_data('records', endpoint='SkaterRecords', team_name=original_six)

# Create a contingency table of original six team players by position
table(origsix_skate$franchiseName, origsix_skate$positionCode) 
```
_The above table is a contingency table with the counts of players from each team in the data set by their position._

```{r}
# Create a contingency table of original six players that are classified as active or retired
table(origsix_skate$franchiseName, origsix_skate$activePlayer)
```
_The above table is a contingency table with counts of players from original six teams that are either active or retired._

## Create numerical summaries for some quantitative variables at each setting of some of your categorical variables
```{r}

```

## Create at least five plots utilizing coloring, grouping, etc. All plots should have nice labels, titles, etc. You should have at least one bar plot, one histogram, one box plot, and one scatterplot. 
```{r}

```

NOTE THAT CODE SHOULD BE SHOWN IN THE FINAL DOCUMENT UNLESS IT IS BEHIND THE SCENES OR UNIMPORTANT FOR THE OUTPUT WE ARE SEEING

