---
title: "R Project 1"
output: 
  github_document: 
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Required Packages
```{r, echo=TRUE, message=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
```

# Contact the NHL records API and the NHL stats API  

## nhl_records function:
```{r}
nhl_records <- function(endpoint, team_id=NULL, team_name=NULL){
  # Endpoint options are: Franchise, TeamTotals, Franchise_SeasonTotals, GoalieRecords, SkaterRecords
  base_url <- 'https://records.nhl.com/site/api'
  if(endpoint == 'Franchise'){
    fork <- '/franchise' 
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url)
  }
  if(endpoint == 'TeamTotals'){
    fork <- '/franchise-team-totals'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url)
  }
  if(endpoint == 'Franchise_SeasonTotals'){
    fork <- '/franchise-season-records'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  if(endpoint == 'GoalieRecords'){
    fork <- '/franchise-goalie-records'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  if(endpoint == 'SkaterRecords'){
    fork <- '/franchise-skater-records' 
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  
  # Convert to text
  df_text <- content(df_get, as='text', enconding='UTF-8')
  
  # Convert form JSON
  df_json <- fromJSON(df_text, flatten=TRUE)
  
  # Use conditional logic to determine how the user is searching
  if(!is.null(team_name) & is.null(team_id)){
    if(endpoint=='Franchise'){
      df <- df_json$data[,-1]
      
      # Clean the dataframe to make the names searchable in our current format
      teamCommonName <- paste(df$teamPlaceName, df$teamCommonName)
      df <- cbind(df, teamCommonName)
    
      # Remove the old names split over two columns
      df <- df[,-(4:5)]
    
      df <- filter(df, teamCommonName==team_name)
    }
    if(endpoint=='TeamTotals'){
      df <- df_json$data[,-1]
      df <- filter(df, teamName==team_name)
    }
    if(endpoint=='Franchise_SeasonTotals' | endpoint=='GoalieRecords' | endpoint=='SkaterRecords'){
      df <- df_json$data[,-1]
      df <- filter(df, franchiseName==team_name)
    }
  }
  if(is.null(team_name) & !is.null(team_id)){
    if(endpoint=='Franchise'){
      df <- df_json$data[,-1]
      df <- filter(df, mostRecentTeamId==as.integer(team_id))
    }
    if(endpoint=='TeamTotals'){
      df <- df_json$data[,-1]
      df <- filter(df, teamId==team_id)
    }
    if(endpoint=='Franchise_SeasonTotals' | endpoint=='GoalieRecords' | endpoint=='SkaterRecords'){
      df <- df_json$data[,-1]
      df <- filter(df, franchiseId==team_id)
    }
  }  
  if(is.null(team_id) & is.null(team_name)){
    df <- df_json$data[,-1]
  }
  return(df)
}
```
```{r, echo=TRUE, eval=FALSE}
nhl_records(endpoint, team_id=NULL, team_name=NULL)
```
The `nhl_records` function pulls data from `records.nhl.com` and outputs it to a data frame. It takes on multiple arguments:  
  
* `endpoint`: can take on five values (`Franchise, TeamTotals, Franchise_SeasonTotals, GoalieRecords, SkaterRecords`). This option will pull data from different API tables. The value for this variable must be passed as a string.  
* `team_id`:  is a unique numeric identifier for each NHL team. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  
* `team_name`: is a unique character assignment for each NHL team. Team names take on the format '<City> <Club Name>'. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  

## nhl_stats function:
```{r}
nhl_stats <- function(modifier, modifier_extension=NULL, team_id=NULL){
  # Modifier options are: team_roster, person_names, team_schedule_next, team_schedule_previous, 
  # team_stats, team_roster_season, team_id, single_season_playoffs
  base_url <- paste0('https://statsapi.web.nhl.com/api/v1/teams/', team_id)
  
  if(modifier=='team_roster'){
    # Pull a GET request
    query <- paste0('?expand=team.roster', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='person_names'){
    # Pull a GET request
    query <- paste0('?expand=person.names', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_schedule_next'){
    # Pull a GET request
    query <- paste0('?expand=team.schedule.next', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_schedule_previous'){
    # Pull a GET request
    query <- paste0('?expand=team.schedule.previous', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_stats'){
    # Pull a GET request
    query <- paste0('?expand=team.stats', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_roster_season'){
    # Pull a GET request
    query <- paste0('?expand=team.roster&season=', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_id'){
    # Pull a GET request
    query <- paste0('?teamId=', modifier_extension)
    df_get <- GET(base_url, query=modifier_full)
  }
  if(modifier=='single_season_playoffs'){
    # Pull a GET request
    query <- paste0('?stats=statsSingleSeasonPlayoffs', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  
  # Convert to text
  df_text <- content(df_get, as='text', enconding='UTF-8')
  
  # Convert form JSON
  df_json <- fromJSON(df_text, flatten=TRUE)
  
  # Remove the redundant ID column
  df <- df_json$teams[,-1]
  
  return(df)
}
```
```{r, echo=TRUE, eval=FALSE}
nhl_stats(modifier, modifier_extension=NULL, team_id=NULL)
```
The `nhl_stats` function pulls data from `www.statsapi.web.nhl.com` and outputs it to a data frame. It takes on multiple arguments:  
  
* `modifier`: can take on one of eight values (`team_roster, person_names, team_schedule_next, team_schedule_previous, team_stats, team_roster_season, team_id, single_season_playoff`). This value will run an expand command on the API. The value for this variable must be passed as a string.  
* `modifier_extension`: Allows the user to pass additional search options to the API. An example of this would be `&gameType=P` to pull data from playoff games only. By default, this value is set to `NULL`. The value for this variable must be passed as a string.  
* `team_id`: is a unique numeric identifier for each NHL team. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  
  
## nhl_data function:
```{r}
nhl_data <- function(source, ...){
  if(source == 'records'){
    return(nhl_records(...))
  }
  if(source == 'stats'){
    return(nhl_stats(...))
  }
}
```
```{r, eval=FALSE, echo=TRUE}
nhl_data(source, ...)
```
The `nhl_data` function serves as a wrapper function for `nhl_records` and `nhl_stats`. 
  
  * `source`: calls whichever function is needed (`records` or `stats`). This argument must be passed a string.  
  * `...`: allows the user to pass all arguments supported by the functions `nhl_records` and `nhl_stats`.

# Basic exploratory data analysis (be sure to discuss what each graph is showing)

## Do a join on two returned datasets from different API endpoints
```{r}

```

## Create at least two new variables
```{r}

```

## Create some contingency tables
```{r}

```

## Create numerical summaries for some quantitative variables at each setting of some of your categorical variables
```{r}

```

## Create at least five plots utilizing coloring, grouping, etc. All plots should have nice labels, titles, etc. You should have at least one bar plot, one histogram, one box plot, and one scatterplot. 
```{r}

```

NOTE THAT CODE SHOULD BE SHOWN IN THE FINAL DOCUMENT UNLESS IT IS BEHIND THE SCENES OR UNIMPORTANT FOR THE OUTPUT WE ARE SEEING

