---
title: "R Project 1"
output: 
  github_document: 
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Required Packages
```{r, echo=TRUE, message=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
library(ggplot2)
```

# Contact the NHL records API and the NHL stats API  

## nhl_records function:
```{r}
nhl_records <- function(endpoint, team_id=NULL, team_name=NULL){
  # Endpoint options are: Franchise, TeamTotals, Franchise_SeasonTotals, GoalieRecords, SkaterRecords
  base_url <- 'https://records.nhl.com/site/api'
  if(endpoint == 'Franchise'){
    fork <- '/franchise' 
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url)
  }
  if(endpoint == 'TeamTotals'){
    fork <- '/franchise-team-totals'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url)
  }
  if(endpoint == 'Franchise_SeasonTotals'){
    fork <- '/franchise-season-records'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  if(endpoint == 'GoalieRecords'){
    fork <- '/franchise-goalie-records'
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  if(endpoint == 'SkaterRecords'){
    fork <- '/franchise-skater-records' 
    
    # Paste together the full url
    full_url <- paste0(base_url, fork)
    
    #Pull a GET request
    df_get <- GET(full_url, query='?cayenneExp=franchiseId=ID')
  }
  
  # Convert to text
  df_text <- content(df_get, as='text', enconding='UTF-8')
  
  # Convert form JSON
  df_json <- fromJSON(df_text, flatten=TRUE)
  
  # Use conditional logic to determine how the user is searching
  if(!is.null(team_name) & is.null(team_id)){
    if(endpoint=='Franchise'){
      df <- df_json$data[,-1]
      
      # Clean the dataframe to make the names searchable in our current format
      teamCommonName <- paste(df$teamPlaceName, df$teamCommonName)
      df <- cbind(df, teamCommonName)
    
      # Remove the old names split over two columns
      df <- df[,-(4:5)]
    
      df <- filter(df, teamCommonName==team_name)
    }
    if(endpoint=='TeamTotals'){
      df <- df_json$data[,-1]
      df <- filter(df, teamName==team_name)
    }
    if(endpoint=='Franchise_SeasonTotals' | endpoint=='GoalieRecords' | endpoint=='SkaterRecords'){
      df <- df_json$data[,-1]
      df <- filter(df, franchiseName==team_name)
    }
  }
  if(is.null(team_name) & !is.null(team_id)){
    if(endpoint=='Franchise'){
      df <- df_json$data[,-1]
      df <- filter(df, mostRecentTeamId==as.integer(team_id))
    }
    if(endpoint=='TeamTotals'){
      df <- df_json$data[,-1]
      df <- filter(df, teamId==team_id)
    }
    if(endpoint=='Franchise_SeasonTotals' | endpoint=='GoalieRecords' | endpoint=='SkaterRecords'){
      df <- df_json$data[,-1]
      df <- filter(df, franchiseId==team_id)
    }
  }  
  if(is.null(team_id) & is.null(team_name)){
    df <- df_json$data[,-1]
  }
  return(df)
}
```
```{r, echo=TRUE, eval=FALSE}
nhl_records(endpoint, team_id=NULL, team_name=NULL)
```
The `nhl_records` function pulls data from `records.nhl.com` and outputs it to a data frame. It takes on multiple arguments:  
  
* `endpoint`: can take on five values (`Franchise, TeamTotals, Franchise_SeasonTotals, GoalieRecords, SkaterRecords`). This option will pull data from different API tables. The value for this variable must be passed as a string.  
* `team_id`:  is a unique numeric identifier for each NHL team. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  
* `team_name`: is a unique character assignment for each NHL team. Team names take on the format '<City> <Club Name>'. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  

## nhl_stats function:
```{r}
nhl_stats <- function(modifier, modifier_extension=NULL, team_id=NULL){
  # Modifier options are: team_roster, person_names, team_schedule_next, team_schedule_previous, 
  # team_stats, team_roster_season, team_id, single_season_playoffs
  base_url <- paste0('https://statsapi.web.nhl.com/api/v1/teams/', team_id)
  
  if(modifier=='team_roster'){
    # Pull a GET request
    query <- paste0('?expand=team.roster', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='person_names'){
    # Pull a GET request
    query <- paste0('?expand=person.names', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_schedule_next'){
    # Pull a GET request
    query <- paste0('?expand=team.schedule.next', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_schedule_previous'){
    # Pull a GET request
    query <- paste0('?expand=team.schedule.previous', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_stats'){
    # Pull a GET request
    query <- paste0('?expand=team.stats', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_roster_season'){
    # Pull a GET request
    query <- paste0('?expand=team.roster&season=', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  if(modifier=='team_id'){
    # Pull a GET request
    query <- paste0('?teamId=', modifier_extension)
    df_get <- GET(base_url, query=modifier_full)
  }
  if(modifier=='single_season_playoffs'){
    # Pull a GET request
    query <- paste0('?stats=statsSingleSeasonPlayoffs', modifier_extension)
    df_get <- GET(base_url, query=query)
  }
  
  # Convert to text
  df_text <- content(df_get, as='text', enconding='UTF-8')
  
  # Convert form JSON
  df_json <- fromJSON(df_text, flatten=TRUE)
  
  # Remove the redundant ID column
  df <- df_json$teams[,-1]
  
  return(df)
}
```
```{r, echo=TRUE, eval=FALSE}
nhl_stats(modifier, modifier_extension=NULL, team_id=NULL)
```
The `nhl_stats` function pulls data from `www.statsapi.web.nhl.com` and outputs it to a data frame. It takes on multiple arguments:  
  
* `modifier`: can take on one of eight values (`team_roster, person_names, team_schedule_next, team_schedule_previous, team_stats, team_roster_season, team_id, single_season_playoff`). This value will run an expand command on the API. The value for this variable must be passed as a string.  
* `modifier_extension`: Allows the user to pass additional search options to the API. An example of this would be `&gameType=P` to pull data from playoff games only. By default, this value is set to `NULL`. The value for this variable must be passed as a string.  
* `team_id`: is a unique numeric identifier for each NHL team. By passing this, the value returned will only contain data for that specific NHL team. By default, this value is set to `NULL`.  
  
## nhl_data function:
```{r}
nhl_data <- function(source, ...){
  if(source == 'records'){
    return(nhl_records(...))
  }
  if(source == 'stats'){
    return(nhl_stats(...))
  }
}
```
```{r, eval=FALSE, echo=TRUE}
nhl_data(source, ...)
```
The `nhl_data` function serves as a wrapper function for `nhl_records` and `nhl_stats`. 
  
  * `source`: calls whichever function is needed (`records` or `stats`). This argument must be passed a string.  
  * `...`: allows the user to pass all arguments supported by the functions `nhl_records` and `nhl_stats`.

# Some exploratory data analysis  

## Join on two returned datasets from different API endpoints
```{r, echo=TRUE, message=FALSE}
# Import the first data set using 'nhl_records' via the wrapper function
dat_1 <- nhl_data('records', endpoint='Franchise', team_name='Boston Bruins')

# Import the second data set using 'nhl_stats' via the wrapper function 
dat_2 <- nhl_data('stats', modifier='team_stats', team_id=6)

# Rename the full team name variable 'name' in 'dat_2' to prepare to merge the data sets
dat_2 <- dplyr::rename(dat_2, teamCommonName=name)

# Merge the two datasets
dat_3 <- merge(dat_1, dat_2, by='teamCommonName')

# Partial output
dat_3[1:5]
```

## Create two new variables  

* New variable: __Average games started__  
This variable will look at the average amount of games started by Carolina Hurricane goaltenders.
```{r, echo=TRUE, message=FALSE}
# Create a new variable, average_games_started
car_goal <- nhl_data('records', endpoint='GoalieRecords', team_name='Carolina Hurricanes')
car_avg_game <- mean(car_goal$gamesPlayed)
average_games_started <- rep(car_avg_game, times=nrow(car_goal))

# Add a new column corresponding to the average number of games a goaltender starts as a Hurricane
car_goal <- cbind(car_goal, average_games_started)

# Display a partial output
head(car_goal[,c(2,7,4,6,29)])
```

* New variable: __Hat trick__  
A hat trick is three or more goals scored by a player in a single game. This variable will determine if a player has scored a hat trick in their career.
```{r, echo=TRUE, message=FALSE}
# Create a new variable, hat_trick
bos_skate <- nhl_data('records', endpoint='SkaterRecords', team_name='Boston Bruins')
hat_trick <- ifelse(bos_skate$mostGoalsOneGame > 2, TRUE, FALSE)

# Add a column corresponding to the new variable hat_trick
bos_skate <- cbind(bos_skate, hat_trick)

# A partial output including our new variable
head(bos_skate[,c(3,9,5,8,15,30)])
```

## Contingency tables
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Import the entire 'SkaterRecords' data frame for the original six teams
original_six <- c('Chicago Blackhawks', 'Boston Bruins', 'Montréal Canadiens', 'Detroit Red Wings', 
                  'New York Rangers', 'Toronto Maple Leafs')
origsix_skate <- nhl_data('records', endpoint='SkaterRecords', team_name=original_six)

# Create a contingency table of original six team players by position
table(origsix_skate$franchiseName, origsix_skate$positionCode) 
```
_The above table is a contingency table with the counts of players from each team in the data set by their position._

```{r}
# Create a contingency table of original six players that are classified as active or retired
table(origsix_skate$franchiseName, origsix_skate$activePlayer)
```
_The above table is a contingency table with counts of players from original six teams that are either active or retired._

## Numerical summaries for some quantitative variables at each setting of some of the categorical variables
```{r, echo=TRUE, message=FALSE}
# Import the full 'SkaterRecords' data set
skater_full <- nhl_data('records', endpoint='SkaterRecords') 
```

```{r, echo=TRUE}
# Mean goals scored by position
round(tapply(skater_full$goals, skater_full$positionCode, mean), 2)
```
_The summary statistics above show mean amount of goals scored by position. As can be seen, centers and wingers typically show up in the scoring column more than defensemen_

```{r, echo=TRUE}
# Mean goals scored by team
round(tapply(skater_full$goals, skater_full$franchiseName, mean), 2)
```
_The summary statistics above show average goals scored by team. As can be seen, franchises which no longer exist seem to be outliers compared to the rest of the data set. This could possibly be a result of the style of play changing over time and not being representative of what scores typically were when they were active franchises._

```{r, echo=TRUE}
# Mean penalty minutes by position
round(tapply(skater_full$penaltyMinutes, skater_full$positionCode, mean), 2)
```
_The summary statistics above show the average penaly minutes by position. Defensemen average the highest amount of penalty minutes, which centers average the least. It might be interesting to try to drill into if strategy or style of play factors into these means._

```{r, echo=TRUE}
# Mean penalty minutes by team
round(tapply(skater_full$penaltyMinutes, skater_full$franchiseName, mean), 2)
```
_The summary statistics above show the average penalty minutes by franchise. Once again we see defunct franchises (i.e. `Montreal Wanderers`) and new franchises (i.e. `Vegas Golden Knights`) do not necessarily fit in with the remainder of the dataset. This would raise questions about deeper trends within the data set._

## Plots utilizing the fetched data and common options for `ggplot` functions.  
```{r, echo=TRUE, message=FALSE}
goalie_records <- dplyr::filter(nhl_data('records', endpoint='GoalieRecords'), activePlayer==TRUE)

# Bar plot
g <- ggplot(data=goalie_records, aes(x=franchiseName))

g + geom_bar(fill='Red') + ggtitle('Active Goaltenders by Franchise ID') + xlab('Franchise Name') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```  

_The above graph shows active NHL goaltenders in the NHL by the franchise name that they play for. Defunct teams, which have no active players, have been removed from the plot._  

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Histogram
franchise_seasontotals <- nhl_data('records', endpoint='Franchise_SeasonTotals')

g <- ggplot(data=franchise_seasontotals, aes(x=fewestLosses))

g + geom_histogram(fill='Dark Orange') + xlab('Fewest Losses') + ylab('Count') + 
  ggtitle('Count of Season with Fewest Losses')
```  

_The above plot shows a histogram with the count of the fewest losses by each NHL team._

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Box plot
goalie_records <- nhl_data('records', endpoint='GoalieRecords')

g <- ggplot(data=goalie_records, aes(x=as.factor(franchiseName), y=wins))

g + geom_boxplot(fill='Light Blue') + ylab('Wins by Goalie') + xlab('Franchise Name') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle('Goalie Wins by Team')
```  

_The above plot is a box-plot of goalie wins by the franchise that they played for. Marty Brodeur is the obvious outlier for the New Jersey Devils._

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Scatterplot
g <- ggplot(data=franchise_seasontotals, aes(x=franchiseId, y=homeWinStreak))

g + geom_point() + ggtitle('Home Win Streak Count by Franchise ID') + xlab('Franchise ID') + ylab('Home Win Streak')
```  

_The above scatterplot shows Home Win Streak counts by Franchise ID. Franchise ID 6, 12, and 16 appear to be outliers. Formal testing could be used to determine if they are actually different from the mean_

```{r}
# Scatterplot
g <- ggplot(data=bos_skate, aes(x=gamesPlayed, y=goals))

g + theme_dark() + geom_point(color='Yellow') + xlab('Games Played') + ylab('Goals Scored') +
  ggtitle('Games Played vs. Goals for Boston Bruins Players')
```  

_The above plot is a scatterplot of games played vs goals scored. A higher y-axis value with a lower x-axis value would suggest higher goal production._
